Reviewer Comments (Scientific Reports)

### 2020-03-28 notebook/sci_rep_reviews.Rmd
Fix line numbers in response to reviewers

I again thank the reviewer for their comments. I agree that it is especially relevant that my manuscript focuses on finite S values, and I have made more emphasis on this point in my revision (e.g., L48, L88, L209, L234). In particular, in addition to the statement on L48, I now state more explicitly in describing the methods that "the objective is to determine the effect of $\gamma$ as system complexity increases toward the boundary at which stability is realistic for a finite system." (L88)

@@

I agree with the relevance of Dougoud et al. (2018), and have made appropriate links to their results on feasibility (L143, L274). Nevertheless, one major result of my manuscript is that varying component response rates (i.e., $\gamma$) only affect stability, not feasibility (L148). Hence, Reviewer 1 is incorrect that my conclusions concerning varying component response rates should be expected to enhance feasibility in experimentally observed food webs (and therefore that stability should be enhanced for biologically realistic ecological communities in which all population densities are positive). Checking for validity in empirical food webs is therefore not entirely appropriate because the consequences of such an investigation would have negligible practical application (but, as I note in the Discussion L194, empirical food webs do not exist in a vacuum, and instead are parts of larger systems within which component response rate variation might indeed be relevant).

I welcome Reviewer 1's call for expanding the scope of models considered for simulations, which was also recommended by Reviewer 2. Nevertheless, my manuscript focuses broadly on complex systems rather than only on pure ecological systems, so I have revised my manuscript to test more general models of common network structures. In terms of ecological models, in addition to random competition, mutualism, and predator-prey networks, I have now also included a cascade food web model (which is a foundation for the extended niche and nested-heirarchy models that Reviewer 1 mentions; Williams and Martinez 2000). Other network types that I have incorporated into my revision include small-world networks (Watts and Strogatz 1998) and scale-free networks (Albert and Barab\'{a}si 2002). Overall, the structured networks that I now include in the main text and supporting information encompass a wide range of physical, biological, and social systems. Results for thse simulations are presented in a new section entitled, "Simulations of structured M across S" (L123). Methods for constructing small-world (L247), scale-free (L260), and cascade food web (L268) networks are now explained, with additional details and figures now in the Supplemental Information. The results from these structured networks match those of the random networks; variation in $\gamma$ increases the number of stable systems given sufficiently high system complexity, thereby further showing the generality of my results across complex systems.

Feasibility and how it relates to ecological systems is explained on L141, with discussion starting on L188 for how this relates to ecological systems. Full details of these results have also been added to the supporting information, and code for building structured networks is publicly available as a documented R package on GitHub (https://github.com/bradduthie/RandomMatrixStability) and accompanying website (https://bradduthie.github.io/RandomMatrixStability).

@@

I have explained the result shown by the black line of Figures 3 (now 4), and the former figure 4 (now in the SI), now more clearly in my revision (L420).

@@

I appreciate these comments, and have tried to provide a more analytical explanation of the changes introduced by \gamma (L53-L64), including a new subsection explaining the effect of variation in $\gamma$ on the (co)variation of M (L65-L81). I agree that these analytical results improve the manuscript, and I think that they are particularly useful in setting the stage for the numerical simulations that now follow. In making these changes, I have slightly adjusted the notation so that the text is more readable.

@@

As suggested, I also considered matrices with even more structure, performing new simulations on each. Specifically, I have simulated three broad classes of structured networks that are widespread in physical, social, and biological systems. These include food webs with cascade structures (e.g., Williams and Martinez 2000), small-world networks (Watts and Strogatz 1998), and scale-free networks (Albert and Barab\'{a}si 2002). Results for these networks are presented in L123-L140, and methods are explained in L247-L273, and in the Supplemental Information. Overall, these structured networks were affected by variation in $\gamma$ in the same way as random networks (increasing stability given sufficiently high complexity).

@@

I thank Reviewer 3 for pointing out these articles. I have cited them where appropriate (L25, L69, L323, L325, and L367).

>>>>

I again thank the reviewer for their comments. I agree that it is especially relevant that my manuscript focuses on finite S values, and I have made more emphasis on this point in my revision (e.g., L48, L88, L207, L232). In particular, in addition to the statement on L48, I now state more explicitly in describing the methods that "the objective is to determine the effect of $\gamma$ as system complexity increases toward the boundary at which stability is realistic for a finite system." (L88)

@@

I agree with the relevance of Dougoud et al. (2018), and have made appropriate links to their results on feasibility (L140, L273). Nevertheless, one major result of my manuscript is that varying component response rates (i.e., $\gamma$) only affect stability, not feasibility (L147). Hence, Reviewer 1 is incorrect that my conclusions concerning varying component response rates should be expected to enhance feasibility in experimentally observed food webs (and therefore that stability should be enhanced for biologically realistic ecological communities in which all population densities are positive). Checking for validity in empirical food webs is therefore not entirely appropriate because the consequences of such an investigation would have negligible practical application (but, as I note in the Discussion L193, empirical food webs do not exist in a vacuum, and instead are parts of larger systems within which component response rate variation might indeed be relevant).

I welcome Reviewer 1's call for expanding the scope of models considered for simulations, which was also recommended by Reviewer 2. Nevertheless, my manuscript focuses broadly on complex systems rather than only on pure ecological systems, so I have revised my manuscript to test more general models of common network structures. In terms of ecological models, in addition to random competition, mutualism, and predator-prey networks, I have now also included a cascade food web model (which is a foundation for the extended niche and nested-heirarchy models that Reviewer 1 mentions; Williams and Martinez 2000). Other network types that I have incorporated into my revision include small-world networks (Watts and Strogatz 1998) and scale-free networks (Albert and Barab\'{a}si 2002). Overall, the structured networks that I now include in the main text and supporting information encompass a wide range of physical, biological, and social systems. Results for thse simulations are presented in a new section entitled, "Simulations of structured M across S" (L122). Methods for constructing small-world (L246), scale-free (L259), and cascade food web (L267) networks are now explained, with additional details and figures now in the Supplemental Information. The results from these structured networks match those of the random networks; variation in $\gamma$ increases the number of stable systems given sufficiently high system complexity, thereby further showing the generality of my results across complex systems.

Feasibility and how it relates to ecological systems is explained on L140, with discussion starting on L187 for how this relates to ecological systems. Full details of these results have also been added to the supporting information, and code for building structured networks is publicly available as a documented R package on GitHub (https://github.com/bradduthie/RandomMatrixStability) and accompanying website (https://bradduthie.github.io/RandomMatrixStability).

@@

I have explained the result shown by the black line of Figures 3 (now 4), and the former figure 4 (now in the SI), now more clearly in my revision (L419).

@@

I appreciate these comments, and have tried to provide a more analytical explanation of the changes introduced by $\gamma$ (L53-L64), including a new subsection explaining the effect of variation in $\gamma$ on the (co)variation of M (L65-L81). I agree that these analytical results improve the manuscript, and I think that they are particularly useful in setting the stage for the numerical simulations that now follow. In making these changes, I have slightly adjusted the notation so that the text is more readable.

@@

As suggested, I also considered matrices with even more structure, performing new simulations on each. Specifically, I have simulated three broad classes of structured networks that are widespread in physical, social, and biological systems. These include food webs with cascade structures (e.g., Williams and Martinez 2000), small-world networks (Watts and Strogatz 1998), and scale-free networks (Albert and Barab\'{a}si 2002). Results for these networks are presented in L122-L139, and methods are explained in L246-L272, and in the Supplemental Information. Overall, these structured networks were affected by variation in $\gamma$ in the same way as random networks (increasing stability given sufficiently high complexity).

@@

I thank Reviewer 3 for pointing out these articles. I have cited them where appropriate (L25, L69, L322, L324, and L366).
